---
title: "Week 4: Visualizing Associations and Models"
author: "Emorie D Beck"
format: 
  revealjs:
    incremental: true
    code-tools: true
    code-copy: true
    code-line-numbers: true
    code-link: true
    preview-links: true
    slide-number: true
    self-contained: true
    css: custom.css
    theme: psc290
    # highlight-style: tango
    margin-left: "0"
    margin-right: "0"
    width: 1400
    # height: 900
    footer: "PSC 290 - Data Visualization"
    logo: "https://github.com/emoriebeck/psc290-data-viz-2022/raw/main/01-week1-intro/02-code/02-images/ucdavis_logo_blue.png"
  # pptx: 
  #   incremental: true  
  #   # code-tools: true
  #   # code-copy: true
  #   # code-line-numbers: true
  #   # code-link: true
  #   css: custom.css
  #   theme: psc290
  #   highlight-style: atom-one-dark
  #   footer: "PSC 290 - Data Visualization"
  #   logo: "https://github.com/emoriebeck/psc290-data-viz-2022/raw/main/01-week1-intro/02-code/02-images/ucdavis_logo_blue.png"
editor_options: 
  chunk_output_type: console
---

```{r, echo = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, error = F)
options(knitr.kable.NA = '')
```

```{r, echo = F}
library(RColorBrewer)
library(knitr)
library(kableExtra)
library(plyr)
library(broom)
library(tidyverse)
```

```{r, eval = F, echo = F}
wd <- "/Volumes/Emorie/other projects/selection"
loadRData <- function(fileName, obj){
#loads an RData file, and returns it
    path <- sprintf("%s/data/sca/%s.RData", wd, fileName)
    load(path)
    get(ls()[grepl(obj, ls())])
}

sample_fun <- function(x){
  x2 <- x %>%
    group_by(study, o_value) %>%
    nest() %>%
    ungroup() %>%
    mutate(data = map(data, 
        ~(.) %>% filter(row_number() %in% sample(1:nrow(.), if(nrow(.) > 500) 500 else nrow(.))))) %>%
    unnest(data) %>%
    select(-Trait, -Outcome)
  x2 <- x2 %>% mutate(study = mapvalues(study, unique(study), paste0("Study", 1:length(unique(study)))))
}

pred_data <- crossing(
  Trait = c("E", "A", "C", "N", "O")
  , Outcome = "mortality"
  , Moderator = c("none", "gender")
) %>%
  mutate(file = paste("sca", Trait, Outcome, sep = "_")
         , data = map2(file, "df1", loadRData)
         , data = map(data, sample_fun)) %>%
  select(-file) 
pred_data <- pred_data$data[[4]]

save(pred_data, file = "/Volumes/Emorie/GitHub/psc290-data-viz-2022/04-week4-associations/04-data/week4-data.RData")
```

# Review
## Review  
* Last week, we talked about how to visualize proportions and probabilities, which often assumes the conjunction of categorical and count data
* This week, we'll focus not on relative differences across categories but rather on how measures co-vary
* There's a ton of ways to do this statistically, which is not the focus of this course or this workshop
* Instead, we'll focus on visualizations that work across cases (e.g., correlations) and model types (lots of generalizable support in `R` for different modeling packages!)

# Final Project Proposals

## Final Project Proposals
* Due at 11:59 PM PST on October 31, 2022
* 1-2 page (single spaced) proposal
  + Short background (why do you / we care?)
  + Research question(s; only the question[s] you are focusing on)
  + Short method (what are the data? Which are you using?)
  + Visualization plan (short summary of your proposed visualization)
  + Challenges and barriers (what do you struggle with this visualization; are there specific barriers?)

## Final Project Proposals  {.smaller}
### Visualization Plan

Can be any combination of the following: 

* Text description
  + clearly describe axes/scales, fills/colors, panels, etc. 
  + describe affordances you will build into the visualization to aid understanding
  + describe why you chose this visualization 
* Draft visualization (digital)
  + rough visualization of some or all of data
* Draft visualization (drawn)
  + some roughly drawn visualization showing what you are hoping to do
  + particularly helpful if you foresee barriers or don't know how to do it
  + can also draw these on a plot of yours from the past, etc. 

## Final Project Proposals  

* The goal of this proposal is to: 
  + Provide you with support to build visualizations you may be nervous about
  + Anticipate barriers and provide you tools
  + Set you up to create a visualization that you are proud of

## Final Project Proposals  

* What is okay? 
  + Introducing a new kind of visualization to your area of research
  + Offering improvements to "standard practice" visualizations in your area
  + Setting up a procedure for a kind of visualization you've long wanted / needed to figure out
  + Really anything that is both useful to you and displays some sort of mastery of course content

# Part 1 Visualizing Associations Among Quantitative Variables

## The Data  
We're going to work with a subsample of data from multiple studies. These are simulated versions of data from my dissertation simulated to mirror the underlying structure of each study, including their drawbacks. 

```{r}
load(url("https://github.com/emoriebeck/psc290-data-viz-2022/blob/main/04-week4-associations/04-data/week4-data.RData?raw=true"))
pred_data
```

## Scatterplots  
* Scatterplots are pretty ubiquitous 
* From a data visualization standpoint, this makes sense
* Scatterplots
  * show raw data  
  * are common enough that little visualization literacy is needed
  * allow for lots of summaries to be placed atop them
  * this is why they are our entry point for today
  
## Scatterplots - Basics  

:::: {.columns}
::: {.column}

* First, let's peek at our data
* These data are based on some of my [dissertation](https://github.com/emoriebeck/selection/raw/master/Beck_Dissertation_2020.pdf) ([Beck & Jackson, 2022](https://psyarxiv.com/7pg9b/)) work that takes data from lots of people across lots of samples and examines personality trait predictors of a bunch of outcomes across all possible combinations of covariates (i.e. specification curve analysis or multiverse analysis)

::: 
::: {.column}

```{r}
pred_data
```

:::
::::

## Scatterplots - Basics  

* We're going to start by looking at associations between Conscientiousness (`p_value`) and self-rated helath (`SRhealth`)
* Both have been rescaled based on the Percentage of Maximum Possible ([Cohen et al., 1999](https://www.tandfonline.com/doi/pdf/10.1207/S15327906MBR3403_2)) with a range of 0-10 to help model convergence

```{r}
pred_data %>% 
  select(study, SID, p_value, SRhealth)
```

## Scatterplots - Basics  {.smaller}

Let's look at a basic scatterplot of the association:  

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(5,3), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    labs(
      x = "Agreeableness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    labs(
      x = "Agreeableness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::::

## Scatterplots - Basics {.smaller}
Let's look at a basic scatterplot of the association **with a trend line**:

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "5"
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    geom_smooth(method = "lm", size = 3, se = F) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    geom_smooth(method = "lm", size = 3, se = F) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::::

## Scatterplots - Basics  {.smaller}
But we have multiple studies, so we need to separate those out using `facet_wrap()`

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "7"
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2) + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::::

## Scatterplots - Basics  {.smaller}
Let's add a trend line again and change the alpha of the points to make them stand out a bit less:  

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "|5-6"
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2, alpha = .25) + 
    geom_smooth(method = "lm", size = 2, se = F) + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2, alpha = .25) + 
    geom_smooth(method = "lm", size = 2, se = F) + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
    ) + 
    theme_classic()
```

:::
::::

## Scatterplots - Basics  {.smaller}
We usually want to have some interval estimate around any average or other measure of central tendency, so we'll set `se = T` in `geom_smooth()`

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "6"
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2, alpha = .25) + 
    geom_smooth(method = "lm", size = 1.5, se = T, color = "black") + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness -- Self-Rated Health Associations Across Samples"
    ) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
pred_data %>% 
  select(study, SID, p_value, SRhealth) %>%
  filter(!is.na(SRhealth)) %>%
  ggplot(aes(x = p_value, y = SRhealth)) + 
    geom_point(shape = 21, fill = "grey80", color = "black", size = 2, alpha = .25) + 
    geom_smooth(method = "lm", size = 1.5, se = T, color = "black") + 
    scale_fill_manual(values = c("grey80", "seagreen4")) + 
    facet_wrap(~study) +
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness -- Self-Rated Health Associations Across Samples"
    ) + 
    theme_classic()
```

:::
::::

## Correlations and Correlelograms  

* Understanding associations is always important, but perhaps never more so than when we do descriptives
* My hot take is that zero-order correlation matrices should always be included in papers
  + Someone's meta-analysis will thank you
* If you're dumping correlations in supplementary materials, then tables are fine
* But you (and your brain) will thank yourself if you use heat maps or correlelograms to visualize the correlations
  + (Remember how quickly and preattentively we perceive color and size?)
* There are `R` packages for this (e.g., [`corrplot`](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)), but where's the fun in that?

## Correlations and Correlelograms  {.smaller}
All right, let's estimate some correlation matrices for each sample:

:::: {.fragment}
:::: {.columns}
::: {.column}

```{r, eval = F}
#| code-line-numbers: "|1-5|6|7-9|10"
r_data <- pred_data %>%
  select(
    study, p_value, age, gender, SRhealth, smokes, exercise
    , BMI, education, parEdu, mortality = o_value
    ) %>%
  mutate_if(is.factor, ~as.numeric(as.character(.))) %>%
  group_by(study) %>%
  nest() %>%
  ungroup() %>%
  mutate(r = map(data, ~cor(., use = "pairwise")))
r_data
```

:::
::: {.column}

```{r, echo = F}
r_data <- pred_data %>%
  select(study, p_value, age, gender, SRhealth, smokes, exercise, BMI, education, parEdu, mortality = o_value) %>%
  mutate_if(is.factor, ~as.numeric(as.character(.))) %>%
  group_by(study) %>%
  nest() %>%
  ungroup() %>%
  mutate(r = map(data, ~cor(., use = "pairwise")))
r_data
```

:::
::::
::::

## Correlations and Correlelograms  

The thing is that we know ggplot doesn't like wide form data, which is what `cor()` produces

```{r}
r_data$r[[1]]
```

## Correlations and Correlelograms  {.smaller}

:::: {.columns}
::: {.column}

* So we need to reshape it in long form
* We're going to use a function so we only have to write the code once and can apply it to all the samples
* Here's what we'll do: 
  + remove the lower triangle and the diagonal of the correlation matrix
  + make matrix a data frame 
  + make the row names of the matrix a column
  + make the columns long
  + factor them to retain order

:::
::: {.column}

:::: {.fragment}

```{r}
#| code-line-numbers: "|4|5|6|7-11|2,12"
r_reshape_fun <- function(r){
  coln <- colnames(r)
  # remove lower tri and diagonal
  r[lower.tri(r, diag = T)] <- NA
  r %>% data.frame() %>%
    rownames_to_column("V1") %>%
    pivot_longer(
      cols = -V1
      , values_to = "r"
      , names_to = "V2"
    ) %>%
    mutate_at(vars(V1, V2), ~factor(., coln))
}

r_data <- r_data %>%
  mutate(r = map(r, r_reshape_fun))
r_data$r[[1]]
```

::::
:::
::::

## Correlations and Correlelograms {.smaller}
### Heat Map Time!
This is, technically, a heat map, but I think we can do better!

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "|2-6|7"
r_data$r[[1]] %>%
  ggplot(aes(
    x = V1
    , y = V2
    , fill = r
  )) + 
  geom_raster() + 
  theme_minimal()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
r_data$r[[1]] %>%
  ggplot(aes(
    x = V1
    , y = V2
    , fill = r
  )) + 
  geom_raster() + 
  theme_minimal()
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Heat Map Time! Colors

Let's add some intuitive colors using `scale_fill_gradient2()`

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "4-11|7-10"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(
    limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue"
    , high = "red"
    , mid = "white"
    , na.value = "white"
    ) + 
  theme_minimal()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(
    limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue"
    , high = "red"
    , mid = "white"
    , na.value = "white"
    ) + 
  theme_minimal()
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Heat Map Time! Labels

Do we need axis labels? Not really -- let's remove them and add `fill` and `title` labels:

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "8-13"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables in Sample 1"
    ) + 
  theme_minimal()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables in Sample 1"
    ) + 
  theme_minimal()
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Heat Map Time! Theme Elements

Let's fix the theme elements. So close!

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "16-23"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(5,5.7), out.width="70%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Heat Map Time! Finishing Touches!

And add text to the correlations using `geom_text()`:

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "4"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  geom_text(aes(label = round(r, 2))) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(5,5.7), out.width="70%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r)) + 
  geom_raster() + 
  geom_text(aes(label = round(r, 2))) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order Correlation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Correlelogram
A correlelogram is basically a heat map that uses size in addition to color.

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, color = r, size = abs(r))) + 
  geom_point() + 
  theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, color = r, size = abs(r))) + 
  geom_point() + 
  theme_classic()
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Correlelogram
We're going to skip the steps we took with a heat map.  
So close! Just need to get rid of that size legend.

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "3,8"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r, size = abs(r))) + 
  geom_point(shape = 21) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  scale_size_continuous(range = c(3,14)) + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order\nCorrelation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(5,5.5), out.width="70%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r, size = abs(r))) + 
  geom_point(shape = 21) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  scale_size_continuous(range = c(3,14)) + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order\nCorrelation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::::

## Correlations and Correlelograms {.smaller}
### Correlelogram

:::: {.columns}
::: {.column}

We're going to skip the steps we took with a heat map.  
So close! Just need to get rid of that size legend.  

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "16"
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r, size = abs(r))) + 
  geom_point(shape = 21) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  scale_size_continuous(range = c(3,14)) + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order\nCorrelation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  guides(size = "none") + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::: {.column}

To do this, we'll use the `guides()` function!  

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(5,5.5), out.width="70%"}
r_data$r[[1]] %>%
  ggplot(aes(x = V1, y = V2, fill = r, size = abs(r))) + 
  geom_point(shape = 21) + 
  scale_fill_gradient2(limits = c(-1,1)
    , breaks = c(-1, -.5, 0, .5, 1)
    , low = "blue", high = "red"
    , mid = "white", na.value = "white") + 
  scale_size_continuous(range = c(3,14)) + 
  labs(
    x = NULL
    , y = NULL
    , fill = "Zero-Order\nCorrelation"
    , title = "Zero-Order Correlations Among Variables"
    , subtitle = "Sample 1"
    ) + 
  guides(size = "none") + 
  theme_classic() + 
  theme(
    legend.position = "bottom"
    , axis.text = element_text(face = "bold")
    , axis.text.x = element_text(angle = 45, hjust = 1)
    , plot.title = element_text(face = "bold", hjust = .5)
    , plot.subtitle = element_text(face = "italic", hjust = .5)
    , panel.background = element_rect(color = "black", size = 1)
  )
```

:::
::::

# Part 2 Visualizing Associations, Parameters, and Predictions from Models

## Part 2 Visualizing Associations, Parameters, and Predictions from Models {.smaller}

* The goal of data visualization is to tell a story that tables, words, etc. either can't or can't do simply
* Data visualizations aims to clarify complex patterns in data
* Thus far, we've mostly focused on building models from raw data or descriptives of raw data
* But in most research, we lean on inferential statistics and hypothesis testing (frequent or Bayesian) to tell our story
* So next, we'll talk about how to use data visualization to tell stories *with* models

## Part 2 Visualizing Associations, Parameters, and Predictions from Models {.smaller}
### Telling Stories with Models 

* The reality is that there is no generalizable way to do this
* So we will focus on models for which we are interested in specific parameters and/or parameterized our questions
* Why? These have some shared functions across *lots* of packages in R
* For models that don't, that's a data cleaning problem, not a visualization problem
  
## Part 2 Visualizing Associations, Parameters, and Predictions from Models {.smaller}

:::: {.columns}
::: {.column}

* Let's start with a basic model and predict later all-cause mortality from Conscientiousness in Sample 1.  
* The basic form of the model is: 

:::: {.fragment}

$$
logit(\frac{\pi}{1-\pi}) = b_0 + b_1*C_{ij} + \epsilon_{ij}
$$
::::
:::
::: {.column}

:::: {.fragment}

```{r}
ds1 <- pred_data %>% filter(study == "Study1")
m1 <- glm(
  o_value ~ p_value
  , data = ds1
  , family = binomial(link = "logit")
  )
summary(m1)
```  

::::

:::
::::

## Part 2 Visualizing Associations, Parameters, and Predictions from Models {.smaller}

:::: {.columns}
::: {.column}

* Models and other objects in `R` are stored in lists or list-like objects
* We can explore these lots of ways, but one good one is with `str()`

:::
::: {.column}
:::: {.fragment}

```{r}
str(m1)
```

::::
:::
::::

## Models + `broom`
* The `broom` package is great for working with models (and the `broom.mixed` add-on makes it even better)
* We're going to talk about how three its functions can be used for / improve data visualization: 
  + `tidy()`
  + `glance()`
  + `augment()`

## Models + `broom::tidy()`
* Outside of `dplyr`/`tidyr`, `tidy()` is a close contender with `purrr::map()` functions as my most used function
* Why? 
  + When you run a model, base `R` provides the `summary()`, `coef()`, etc. to extract various components of the model
  + But these aren't `data.frames`, which are core input to a lot of other `R` functions across packages
  + `tidy()` provides a data frame with core model coefficients, inferential tests, etc. that be easily matched and merged across models, etc. 

## Models + `broom::tidy()`
* But with logistic regression with a logit link, we are left with coefficents that have to be interpreted in log odds, which realistically, almost no one can do
* So we have to "undo" the log, which you may remember can done by exponentiating the natural log (ln)
* But we can directly exponentiate from the summary function because it's the wrong class of object
* We could just exponentiate the coefficients from the `coef()` function, but this still leaves us with the need to extract estimates of precision, like standard errors, confidence intervals, and more. 

:::: {.fragment}

```{r}
coef(m1)
```

::::

## Models + `broom::tidy()`

Enter `broom::tidy()`!

:::: {.fragment}

```{r}
tidy(m1)
```

::::

## Models + `broom::tidy()`  
Enter `broom::tidy()`!  
Even better, we can easily get confidence intervals  

```{r}
tidy(m1, conf.int = T)
```

## Models + `broom::tidy()` {.smaller}

:::: {.columns}
::: {.column}

* We'd rarely make a plot for just two parameters unless we're doing it across groups or samples  
* Watch! Let's make a nested data frame that will hold
  + All the data for each sample
  + A model for each sample
  + The `tidy()` data frame of the parameter estimates for each sample
  
:::
::: {.column}

:::: {.fragment}

```{r}
#| code-line-numbers: "|3-6|8-14|15"
tidy_ci <- function(m) tidy(m, conf.int = T)

nested_m <- pred_data %>%
  group_by(study) %>%
  nest() %>%
  ungroup() %>%
  mutate(
    m = map(data
            , ~glm(
              o_value ~p_value
              , data = .
              , family = binomial(link = "logit")
              )
            )
    , tidy = map(m, tidy_ci)
  )
nested_m
```

::::

:::
::::

## Models + `broom::tidy()`
Now, we'll drop the `data` and `m` columns that we don't need and `unnest()` our `tidy()` data frames

:::: {.fragment}

```{r}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy)
```

::::

## Models + `broom::tidy()` {.smaller}
Now these parameters from multiple models, we may want to plot!

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "|4|5-7|8-13|14"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    theme_classic()
```

:::
::::

## Models + `broom::tidy()` {.smaller}
Almost, but we have two parameters for each model (Intercept and p_value), so let's split those in a facet:

:::: {.fragment}

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "|14"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    facet_grid(~term) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    facet_grid(~term) + 
    theme_classic()
```

:::
::::
::::

## Models + `broom::tidy()`

We've got some work to do to make this an intuitive figure. Let's: 

  + Add a dashed line at 1 (odd ratio of 1 is a null effect)
  + Make the points bigger
  + Fix the titles on the plot and axis titles
  + Add some color
  + Fiddle with themes to make it prettier

## Models + `broom::tidy()` {.smaller}
Add a dashed line at 1 (odd ratio of 1 is a null effect)

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "8"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point() + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::::

## Models + `broom::tidy()` {.smaller}

Make the points bigger

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "14"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 15) + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 15) + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::::


## Models + `broom::tidy()` {.smaller}

Fix the titles on the plot and axis titles

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "15-21"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 15) + 
    labs(
      x = "Estimate (CI) in OR"
      , y = NULL
      , title = "Conscientiousness was associated with mortality 50% of samples"
      , subtitle = "Samples with lower mortality risk overall had fewer significant associations"
      ) + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 15) + 
    labs(
      x = "Estimate (CI) in OR"
      , y = NULL
      , title = "Conscientiousness was associated with mortality 50% of samples"
      , subtitle = "Samples with lower mortality risk overall had fewer significant associations"
      ) + 
    facet_grid(~term, scales = "free") + 
    theme_classic()
```

:::
::::

## Models + `broom::tidy()` {.smaller}

Add some color  
Fiddle with themes to make it prettier  

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "|6|23-33"
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate, fill = study)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 22) + 
    labs(
      x = "Estimate (CI) in OR"
      , y = NULL
      , title = "Conscientiousness was associated with mortality 50% of samples"
      , subtitle = "Samples with lower mortality risk overall had fewer significant associations"
      ) + 
    facet_grid(~term, scales = "free") + 
    theme_classic() + 
    theme(
      legend.position = "none"
      , axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.2))
      , axis.line = element_blank()
      , strip.text = element_text(face = "bold", size = rel(1.1), color = "white")
      , strip.background = element_rect(fill = "black")
      , plot.title = element_text(face = "bold", size = rel(1.1), hjust = .5)
      , plot.subtitle = element_text(face = "italic", size = rel(1.1))
      , panel.border = element_rect(color = "black", fill = NA, size = 1)
    )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_m %>%
  select(study, tidy) %>%
  unnest(tidy) %>%
  mutate_at(vars(estimate, conf.low, conf.high), exp) %>%
  ggplot(
    aes(y = study, x = estimate, fill = study)
  ) + 
    geom_vline(aes(xintercept = 1), linetype = "dashed") + 
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high)
      , position = position_dodge(width = .9)
      , width = .1
      ) + 
    geom_point(size = 3, shape = 22) + 
    labs(
      x = "Estimate (CI) in OR"
      , y = NULL
      , title = "Conscientiousness was associated with mortality 50% of samples"
      , subtitle = "Samples with lower mortality risk overall had fewer significant associations"
      ) + 
    facet_grid(~term, scales = "free") + 
    theme_classic() + 
    theme(
      legend.position = "none"
      , axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.2))
      , axis.line = element_blank()
      , strip.text = element_text(face = "bold", size = rel(1.1), color = "white")
      , strip.background = element_rect(fill = "black")
      , plot.title = element_text(face = "bold", size = rel(1.1), hjust = .5)
      , plot.subtitle = element_text(face = "italic", size = rel(1.1))
      , panel.border = element_rect(color = "black", fill = NA, size = 1)
    )
```

:::
::::

## Models + `broom::tidy()` 
* This isn't perfect. But we're going to come back to this kind of plot when we talk about "piecing plots together." 
* Personally, I would: 
  + Add text with Est. (CI) and N for each sample in the figure
  + Build both of these separately in order to order by effect size
  + Then put them back together and re-add the title

## Models + `broom`: `glance()` {.smaller}

:::: {.columns}
::: {.column}

* When we run models, we need to care about more that just point and interval estimates
* Often we are interested in comparing models, checking diagnostics, etc. 
* Again, all of these are embedded (mostly), in the model objects 
* The `glance()` function brings some of these important ones into a single object

:::
::: {.column}
:::: {.fragment}

* Here's what it gives us for our logistic regression model

```{r}
glance(m1)
```

::::
:::
::::

## Models + `broom`: `glance()`
* Let's also look for al linear model, which may be more familiar for many of you: 

:::: {.fragment}

```{r}
m2 <- lm(SRhealth ~ age, data = ds1)
glance(m2)
```

::::

## Models + `broom`: `glance()`
As before, we can do this with lots of models to compare across samples: 

```{r}
nested_m 
```

## Models + `broom`: `glance()`
As before, we can do this with lots of models to compare across samples: 

```{r}
#| code-line-numbers: "2"
nested_m <- nested_m %>%
  mutate(glance = map(m, glance))
nested_m
```

## Models + `broom`: `glance()`
As before, we can do this with lots of models to compare across samples  
*Now unnesting:*

:::: {.columns}
::: {.column}

```{r, eval = F}
nested_m %>%
  select(study, glance) %>%
  unnest(glance)
```

:::
::: {.column}

```{r, echo = F}
nested_m %>%
  select(study, glance) %>%
  unnest(glance)
```

:::
::::

## Models + `broom`: `glance()` 
Realistically, this is the kind of info we table, but we can also merge it with info from tidy: 

:::: {.columns}
::: {.column}

```{r, eval = F}
nested_m %>%
  select(-data, -m) %>%
  unnest(tidy) %>% 
  unnest(glance) %>%
  mutate_if(is.numeric, ~round(., 2))
```

:::
::: {.column}

```{r, echo = F}
nested_m %>%
  select(-data, -m) %>%
  unnest(tidy) %>% 
  unnest(glance) %>%
  mutate_if(is.numeric, ~round(., 2)) %>%
  kable(., "html") %>%
  kable_classic(full_width = F, font_size = 20)
```

:::
::::

## Models + `broom`: `augment()` {.smaller}

:::: {.columns}
::: {.column}

* Diagnostics are not just summary statistics! 
* We care a lot about prediction, too
  + Residuals both tell us unexplained variance (i.e. how observed data deviate from model predictions) and how appropriate our model was
  + Model predictions and prediction intervals tell us about how our model is doing across levels our variables

:::
::: {.column}
:::: {.fragment}

Let's keep working with our nested data frame. Remember, it looks like this: 
  
```{r}
nested_m
```

::::
:::
::::

## Models + `broom`: `augment()` {.smaller}

:::: {.columns}
::: {.column width="40%"}

* `augment()` let's us add (augment) the raw data we feed the model based on the fitted model
* Notice we now have more columns  

:::
::: {.column width="60%"}

```{r}
nested_m <- nested_m %>%
  mutate(data = map2(m, data, augment, se_fit = T))
nested_m
```

:::
::::

## Models + `broom`: `augment()` {.smaller}

:::: {.columns}
::: {.column width="40%"}

* Here's the columns we used along with the additional columns with a `glm`: 
  + `.fitted`: fitted / predicted value
  + `.se.fit`: standard error 
  + `.resid`: observed - fitted
  + `.std.resd`: standardized residuals
  + `.sigma`: estimated residual SD when this obs is dropped from model
  + `cooksd`: Cooks distance (is this an outlier?)
  
:::
::: {.column width="60%"}

```{r}
nested_m$data[[1]] %>%
  select(o_value, SID, p_value, .fitted:.cooksd)
```

:::
::::

## Models + `broom`: `augment()` {.smaller}
For the most part, many of the checks with `glm`'s and `lm`'s are the same. But it's a bit easier to wrap your head around `lm()`, so let's switch to that: 

:::: {.columns}
::: {.column}

```{r, eval = F}
#| code-line-numbers: "1-6|7|8|9|10"
nested_lm <- pred_data %>%
  select(study, SID, p_value, age, SRhealth) %>%
  drop_na() %>%
  group_by(study) %>%
  nest() %>%
  ungroup() %>%
  mutate(m = map(data, ~lm(SRhealth ~ p_value + age, data = .))
         , tidy = map(m, tidy_ci)
         , glance = map(m, glance)
         , data = map2(m, data, augment, se_fit = T, interval = "confidence"))
nested_lm
```

:::
::: {.column}

```{r, echo = F}
nested_lm <- pred_data %>%
  select(study, SID, p_value, age, SRhealth) %>%
  drop_na() %>%
  group_by(study) %>%
  nest() %>%
  ungroup() %>%
  mutate(m = map(data, ~lm(SRhealth ~ p_value + age, data = .))
         , tidy = map(m, tidy_ci)
         , glance = map(m, glance)
         , data = map2(m, data, augment, se_fit = T, interval = "confidence"))
nested_lm
```

:::
::::

## Models + `broom`: `augment()` {.smaller}
:::: {.columns}
::: {.column}

* Here's the columns we used along with the additional columns with an `lm`: 
  + `.fitted`: fitted / predicted value
  + `.se.fit`: standard error
  + `.lower`: lower bound of the confidence/prediction interval
  + `.upper`: upper bound of the confidence/prediction interval
  + `.resid`: observed - fitted
  + `.std.resd`: standardized residuals
  + `.sigma`: estimated residual SD when this obs is dropped from model
  + `cooksd`: Cooks distance (is this an outlier?)

:::
::: {.column}

```{r, eval = F}
nested_lm$data[[1]]
```

```{r, echo=FALSE}
nested_lm$data[[1]] %>%
  kable(., "html", digits = 2) %>%
  kable_classic(full_width = F, font_size = 20) %>%
  scroll_box(height = "500px")
```

:::
::::

## Models + `broom`: `augment()` {.smaller}
* One standard diagnostic plot is to plot fitted values v residuals
* Looks a little wonky (remember, these are results from multiple harmonized studies)

:::: {.fragment}
:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = .fitted
    , y = .resid
  )) + 
  geom_point() + 
  theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = .fitted
    , y = .resid
  )) + 
  geom_point() + 
  theme_classic()
```


:::
::::
::::

## Models + `broom`: `augment()` {.smaller}

* One standard diagnostic plot is to plot fitted values v residuals
* Looks a little wonky (remember, these are results from multiple harmonized studies)

:::: {.fragment}
:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = .fitted
    , y = .resid
  )) + 
  geom_point() +
  labs(
    x = "Model Fitted Values"
    , y = "Residual") + 
  facet_wrap(~study) + 
  theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = .fitted
    , y = .resid
  )) + 
  geom_point() +
  labs(
    x = "Model Fitted Values"
    , y = "Residual") + 
  facet_wrap(~study) + 
  theme_classic()
```

:::
::::
::::

## Models + `broom`: `augment()` {.smaller}

Another is raw v. fitted

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = p_value
    , y = .resid
  )) + 
  geom_point() +
  facet_wrap(~study) + 
  theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
nested_lm %>%
  select(study, data) %>%
  unnest(data) %>%
  ggplot(aes(
    x = p_value
    , y = .resid
  )) + 
  geom_point() +
  facet_wrap(~study) + 
  theme_classic()
```

:::
::::

## Model Predictions {.smaller}
* Although we can get the standard error of the prediction for each person, we often want to look at theoretical predictions, adjusting for covariates. We can typically use built-in `predict()` or `fitted()` functions
* To do this, we need to see theoretical ranges of key variables and grab averages of covariates
* I use functions for this. We'll do one without a function and build there

:::: {.fragment}
:::: {.columns}
::: {.column}

```{r, eval = F}
#| code-line-numbers: "|1|2|4|4-7|10|8-11"
m1 <- nested_lm$m[[1]]
d1 <- m1$model

crossing(
  p_value = seq(0, 10, length.out = 100)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  )
```

:::
::: {.column}

```{r, echo = F}
m1 <- nested_lm$m[[1]]
d1 <- m1$model

crossing(
  p_value = seq(0, 10, length.out = 100)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  )
```

:::
::::
::::

## Model Predictions {.smaller}
### Prediction Interval
:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "9-12|10|11"
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    theme_classic()
```

:::
::::

## Model Predictions {.smaller}
### Confidence Interval
:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "9-12|10|11"
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "confidence")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "confidence")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    theme_classic()
```

:::
::::

## Model Predictions
* This is fine, but it could use some improvements: 
  + better scales
  + raw data 
  + the usual aesthetics
  
## Model Predictions {.smaller}
### Better scales

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "12-13"
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(
    .
    , predict(m1, newdata = ., interval = "prediction")
  ) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    theme_classic()
```

:::
::::

## Model Predictions {.smaller}
### Raw Data 

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "7-12"
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(., predict(m1, newdata = ., interval = "prediction")) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(
      data = d1
      , aes(x = p_value, y = SRhealth)
      , alpha = .4
      , color = "seagreen4"
      ) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    theme_classic()
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(., predict(m1, newdata = ., interval = "prediction")) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(
      data = d1
      , aes(x = p_value, y = SRhealth)
      , alpha = .4
      , color = "seagreen4"
      ) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    theme_classic()
```

:::
::::

## Model Predictions {.smaller}
### The usual aesthetics

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,4), out.width="100%"}
#| code-line-numbers: "13-17, 19-23"
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(., predict(m1, newdata = ., interval = "prediction")) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .4, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated"
      ) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,4), out.width="100%"}
crossing(
  p_value = seq(0, 10, .1)
  , age = mean(d1$age)
) %>%
  bind_cols(., predict(m1, newdata = ., interval = "prediction")) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .4, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated"
      ) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      )
```

:::
::::

## Model Predictions {.smaller}
### Across All Samples

:::: {.columns}
::: {.column}

:::: {.fragment}
```{r}
pred_fun <- function(m){
  d <- m$model

  crossing(
    p_value = seq(0, 10, length.out = 100)
    , age = mean(d$age)
  ) %>%
    bind_cols(
      .
      , predict(m, newdata = ., interval = "prediction")
    )
}
```

::::

:::
::: {.column}
:::: {.fragment}

```{r}
nested_lm <- nested_lm %>%
  mutate(pred = map(m, pred_fun))
nested_lm
```

::::

:::
::::

## Model Predictions
### Across All Samples

```{r}
nested_lm %>%
  select(study, pred) %>%
  unnest(pred)
```

## Model Predictions {.smaller}  

Very close, but our intervals are cut off

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "|16"
nested_lm %>%
  select(study, pred) %>%
  unnest(pred) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .2, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated In Most Samples"
      ) + 
    facet_wrap(~study, ncol = 2) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      , strip.background = element_rect(fill = "darkseagreen4")
      , strip.text = element_text(face = "bold", color = "white")
      )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,6), out.width="100%"}
nested_lm %>%
  select(study, pred) %>%
  unnest(pred) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .2, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated In Most Samples"
      ) + 
    facet_wrap(~study, ncol = 2) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      , strip.background = element_rect(fill = "darkseagreen4")
      , strip.text = element_text(face = "bold", color = "white")
      )
```

:::
::::

## Model Predictions {.smaller}

:::: {.columns}
::: {.column}

```{r, echo = T, eval = F, fig.align='center', fig.dim= c(6,6), out.width="100%"}
#| code-line-numbers: "4-5"
nested_lm %>%
  select(study, pred) %>%
  unnest(pred) %>%
  mutate(upr = ifelse(upr > 10, 10, upr)
         , lwr = ifelse(lwr < 0, 0, lwr)) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .2, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated In Most Samples"
      ) + 
    facet_wrap(~study, ncol = 2) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      , strip.background = element_rect(fill = "darkseagreen4")
      , strip.text = element_text(face = "bold", color = "white")
      )
```

:::
::: {.column}

```{r, echo = F, fig.show='show', fig.align='center', fig.dim= c(6,6), out.width="100%"}
nested_lm %>%
  select(study, pred) %>%
  unnest(pred) %>%
  mutate(upr = ifelse(upr > 10, 10, upr)
         , lwr = ifelse(lwr < 0, 0, lwr)) %>%
  ggplot(aes(x = p_value, y = fit)) + 
    geom_point(data = d1, aes(x = p_value, y = SRhealth)
      , alpha = .2, color = "seagreen4") + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "seagreen4", alpha = .2) + 
    geom_line(color = "seagreen4", size = 2) + 
    scale_x_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    scale_y_continuous(limits = c(0,10.2), breaks = seq(0,10,2)) + 
    labs(
      x = "Conscientiousness (POMP; 0-10)"
      , y = "Predicted Self-Rated Health (POMP; 0-10)"
      , title = "Conscientiousness and Self-Rated Health\nWere Weakly Associated In Most Samples"
      ) + 
    facet_wrap(~study, ncol = 2) + 
    theme_classic() + 
    theme(
      axis.text = element_text(face = "bold", size = rel(1.1))
      , axis.title = element_text(face = "bold", size = rel(1.1))
      , plot.title = element_text(face = "bold", size = rel(1.2), hjust = .5)
      , strip.background = element_rect(fill = "darkseagreen4")
      , strip.text = element_text(face = "bold", color = "white")
      )
```

:::
::::

## Wrapping Up
* This is a quick introduction to visualizing associations and working with models
* Here, we focused on doing things very manually to promote understanding
* But there are lots of packages to automate much of this: 
  + [`sjplot`](https://strengejacke.github.io/sjPlot/)
  + [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/articles/marginaleffects.html)
  + etc.

